### 2022-04-16 paper/Chapters/Hardware.tex

@@

@@

Instead, the first board was built out in a piecewise fashion, allowing each section of the board a way to fail without interrupting the rest of the board.
This allows each function of the board to be tested individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
\todoquestion{This is that tricky sentence}For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, a complex circuit would have to be built that managed Lithium Polymer (LiPo) batteries which could cause fires if mishandled.
Since power is obviously a critical component of the board, A secondary way of feeding power directly the board was built to bypass all that circuitry in case it fails.
In doing so, the board can still be powered and it's other features can be tested while a fix for that portion of the circuitry is developed for the next iteration.

@@

Finally, everything must connect to the CM4.
Figure \ref{fig:CM4Circuit} shows the the two mezzanine connectors the CM4 uses to communicate with the rest of the PCB (schematic in Figure \ref{fig:pcb_schematic_cm4}).
Remember that there is only 0.04mm of space in between each of the pins, leaving very little space to route the traces where they need to go.
The first thing to notice is the large stretch of traces on the back side of the board coming from the HDMI ports.
Because most of the traces come in positive and negative pairs, called differential pairs, the two tracks must be roughly the same length as each other.
This is because as data is transmitted across the pair of tracks, they need to arrive at their destination at the same time based on the physical distance the signals need to cover.
Since the speed of electricity is constant throughout the PCB board, the traces may need to wind back and forth so that the lengths of the positive and negative tracks are equal.
Also note the blank area surrounded by vias at the top of Figure \ref{fig:CM4Circuit}.
This is where the WiFi and Bluetooth antennas are located on the CM4 board.
By removing the copper pour from this area on the PCB board and stitching the perimeter with vias, it prevents the board from acting like another antenna and interfering with the signals being sent to and from the CM4.
Finally, each of the other components are connected to the CM4 board in the same manner as described throughout this chapter, such as the Ethernet adapter, the SD card reader, and remaining ports.
\todoadvisor{Why are those components not described in the same way as power/USB/etc.?}\todoquestion{The Ethernet port is hooked directly into the CM4 connector with nothing interested except that they are also differential pairs, as described in the HDMI section. I suppose they could be explained fully, but I worry that it wouldn't add any useful information.}
See Appendices \ref{AppendixA} and \ref{AppendixB} for further information.

@@

The second board was assembled mostly successfully, and allowed the first stage of testing to begin.
The only issue apart from some ports being too close together with real-world cables was one component having the wrong footprint on the board.
Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in it's datasheet\todoquestion{Does this need to be cited? Can you cite a datasheet like this?}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
This discrepancy meant that that component could not be soldered, and it's associated circuit would not function properly.
In this case, the component was a part of the battery charging circuit, causing the entire primary power supply to be unusable.
Thankfully, as described in Section \ref{sec:Prototyping}, the secondary method of powering the board could be used to test the remaining features of the board.
After making the necessary fixes, a second round of five PCB boards and a stencil was ordered.
Because there was still enough components to assemble three more boards, and no components needed to be changed, no more components were ordered at this stage.
Once the boards were delivered, the revised board was again sent to be professionally assembled.

>>>>

The entirety of the hardware schematics and designs developed for this project are available on GitHub at: \url{https://github.com/Dylancyclone/thesis/kicad}.

@@

Due to the lack of preliminary studies or research into the topic of target audience for this project, the requirements laid out in this section are based on the author's own requirements and preferences.
The requirements are constructed with the target user being someone who already has a working desktop computer and wishes to leverage it's power remotely without sacrificing the user experience.

@@

Instead, the first board was built out in a piecewise fashion, allowing each section of the board a way to fail without interrupting the rest of the board.
This allows each function of the board to be tested individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, a complex circuit would have to be built that managed Lithium Polymer (LiPo) batteries which could cause fires if mishandled.
Since power is obviously a critical component of the board, A secondary way of feeding power directly the board was built to bypass all that circuitry in case it fails.
In doing so, the board can still be powered and it's other features can be tested while a fix for that portion of the circuitry is developed for the next iteration.

@@

Finally, everything must connect to the CM4.
Figure \ref{fig:CM4Circuit} shows the the two mezzanine connectors the CM4 uses to communicate with the rest of the PCB (schematic in Figure \ref{fig:pcb_schematic_cm4}).
Remember that there is only 0.04mm of space in between each of the pins, leaving very little space to route the traces where they need to go.
The first thing to notice is the large stretch of traces on the back side of the board coming from the HDMI ports.
Because most of the traces come in positive and negative pairs, called differential pairs, the two tracks must be roughly the same length as each other.
This is because as data is transmitted across the pair of tracks, they need to arrive at their destination at the same time based on the physical distance the signals need to cover.
Since the speed of electricity is constant throughout the PCB board, the traces may need to wind back and forth so that the lengths of the positive and negative tracks are equal.
Also note the blank area surrounded by vias at the top of Figure \ref{fig:CM4Circuit}.
This is where the WiFi and Bluetooth antennas are located on the CM4 board.
By removing the copper pour from this area on the PCB board and stitching the perimeter with vias, it prevents the board from acting like another antenna and interfering with the signals being sent to and from the CM4.
Finally, each of the other components are connected to the CM4 board in the same manner as described throughout this chapter, such as the Ethernet adapter, the SD card reader, and remaining ports.
These ports are connected by simply connecting the appropriate pin on the CM4's mezzanine connector to the corresponding pin on the component.
Apart from the concerns brought up by keeping the differential pairs in line and the power carefully managed as with the previous components described in detail, there are no special considerations that need to be made while designing the layout for these components.
See Appendices \ref{AppendixA} and \ref{AppendixB} for further information.

@@

The second board was assembled mostly successfully, and allowed the first stage of testing to begin.
The only issue apart from some ports being too close together with real-world cables was one component having the wrong footprint on the board.
Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in its datasheet \cite{MCP73871-2CCI/ML-ND_datasheet}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
This discrepancy meant that that component could not be soldered, and it's associated circuit would not function properly.
In this case, the component was a part of the battery charging circuit, causing the entire primary power supply to be unusable.
Thankfully, as described in Section \ref{sec:Prototyping}, the secondary method of powering the board could be used to test the remaining features of the board.
After making the necessary fixes, a second round of five PCB boards and a stencil was ordered.
Because there was still enough components to assemble three more boards, and no components needed to be changed, no more components were ordered at this stage.
Once the boards were delivered, the revised board was again sent to be professionally assembled.

### 2022-04-16 paper/Chapters/StateOfTheArt.tex
Feedback from advisor and thesis defense

Though the GameStream technology that powers the Nvidia Shield devices is closed source and only officially compatible with the products Nvidia releases, its enticing power and potential drew a community to reverse engineer the protocol.
In 2013, a group of students at Case Western Reserve University developed Moonlight, an open source implementation of the GameStream protocol \cite{moonlight}.
This open source implementation allows the development of GameStream compatible clients that can run on other devices, from other computers to embedded ARM devices.
This stands as a good starting place to develop the solution described in this paper.

@@

\begin{enumerate}
  \item \textbf{Hardware Feasibility:} \emph{What hardware is needed in order to power a mobile device capable of acting as a client?}
  \item \textbf{Hardware Cost:} \emph{Is it feasible to construct such as device at a lower or comparable cost to existing solutions?}
  \item \textbf{Communication Protocol:} \emph{Does a protocol exist that is efficient enough to stream demanding applications to a client?}
  \item \textbf{Software Application:} \emph{Can a software solution be built to utilize this protocol in a manner that performant enough?}
\end{enumerate}

>>>>

Though the GameStream technology that powers the Nvidia Shield devices is closed source and only officially compatible with the products Nvidia releases, its enticing power and potential drew a community to reverse engineer the protocol.
In 2013, a group of students at Case Western Reserve University developed Moonlight, an open source implementation of the GameStream protocol \cite{moonlight}.
The project is structured as a core library written in \emph{C} with a number of community built clients built on top of it for various platforms.
In essence, it turns the closed-source black-box protocol developed by Nvidia into a usable programming interface.
By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for almost any operating system.
The Moonlight community has already developed fantastic clients for many platforms such as Windows, Mac, Linux, Android, iOS, Web, and even other gaming consoles.
All that is needed for the purposes of this project is to make a fork of the project that tweaks it to best suit the CM4 that drives the hardware and develop a suite of tools to test it's functionality.
This stands as a good starting place to develop the solution described in this paper.

@@

\begin{enumerate}
  \item \textbf{Hardware Feasibility:} \emph{What hardware is needed in order to power a mobile device capable of acting as a client?}
  \item \textbf{Hardware Cost:} \emph{Is it feasible to construct such as device at a lower or comparable cost to existing solutions?}
  \item \textbf{Communication Protocol:} \emph{Does a protocol exist that is efficient enough to stream demanding applications to a client?}
  \item \textbf{Software Application:} \emph{Can a software solution be built to utilize this protocol in a manner that is performant enough?}
\end{enumerate}
